---
title: "R_basics_assignment"
author: "Kyra Reisenfeld"
date: "January 14, 2020"
output: html_document
---

```{r}
setwd("~/Documents")
tgpp <- read.csv("./Quant/Quant_assignments/tgpp.csv", header = TRUE)
head(tgpp)
```
1. What are the names of the columns in this dataset?
The names of the columns are "plot, year, record_id, corner, scale, richness, easting, northing, slope, ph, yrsslb"
2. How many rows and columns does this data file have?
There are 11 columns and 4,080 rows in this data file.
3. What kind of object is each data column? Hint: checkout the function sapply().
```{r}
sapply(tgpp, class)
```
4. What are the values of the the datafile for rows 1, 5, and 8 at columns 3, 
7, and 10
```{r}
tgpp[1, 3]
tgpp[1,7]
tgpp[1,10]
tgpp[5,3]
tgpp[5,7]
tgpp[5,10]
tgpp[8, 3]
tgpp[8,7]
tgpp[8,10]
```
OR
```{r}
df<-c(tgpp[1, 3], tgpp[1,7], tgpp[1,10], tgpp[5,3], tgpp[5,7], tgpp[5,10], tgpp[8, 3], tgpp[8,7], tgpp[8,10])
df
```

5. Create a pdf of the relationship between the variables "scale" and "richness". 
Scale is the area in square meters of the quadrat in which richness was
recorded. Be sure to label your axes clearly, and choose a color you find
pleasing for the points. To get a list of available stock colors use the
function colors(). Also see this link: 
<http://research.stowers-institute.org/efg/R/Color/Chart/index.htm>.
```{r}
p<-plot(log(richness)~log(scale), col="red", data=tgpp)
print(p)
```
6. What happens to your plot when you set the plot argument log equal to 'xy'.
`plot(..., log='xy')`
```{r}
plot(richness~scale, col="red", log='xy',data=tgpp)
```

