```{r}
library(vegan)
library(lattice)
data(BCI)
?BCI
## UTM Coordinates (in metres)
BCI_xy = data.frame(x = rep(seq(625754, 626654, by=100), each=5), 
                    y = rep(seq(1011569,  1011969, by=100), len=50))
plot(BCI_xy)
```
Overall Data (just wanted to see)
```{r}
sr = apply(BCI, 1, function(x) sum(x > 0))
hist(sr)
```
```{r}
plot(BCI_xy, cex = sr/max(sr))
```
```{r}
col_brks = hist(sr, plot=F)$breaks
col_indices = as.numeric(cut(sr, col_brks))
cols = rev(terrain.colors(length(col_brks)))
plot(BCI_xy, cex=2, pch=19, col=cols[col_indices])
```
```{r}
# calculate Euclidean distance between richness and spatial coordinates
sr_dist = dist(sr)
xy_dist = dist(BCI_xy)
```

```{r}
max_dist = max(xy_dist) / 2

# plot result
plot(xy_dist, sr_dist)
abline(lm(sr_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, sr_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)
```
```{r}
# compute correlation
obs_cor = cor(xy_dist, sr_dist)
obs_cor
```
```{r}
# carry out a permutation test for significance:
nperm = 1000
null_cor = obs_cor
for (i in 2:nperm) {
    # shuffle the rows of the spatial coordinates
    tmp_xy = BCI_xy[sample(nrow(BCI_xy)), ]
    # correlation between the shuffled spatial coordinates and sr_dist
    null_cor[i] = cor(dist(tmp_xy), sr_dist)
}
# compute the p-value
sum(null_cor >= obs_cor) / nperm 
```
```{r}
# carry out the same analysis
sr_mantel = mantel(xy_dist, sr_dist)
sr_mantel
```
1. Common  
```{r}
alseis = BCI[,"Alseis.blackiana"]
alseis
```

```{r}
plot(BCI_xy, cex = alseis/max(alseis))
```
```{r}
##ggplot solution
library(ggplot2)
ggplot(BCI_xy, aes(x=x, y=y, size=alseis)) + geom_point() + scale_size_continuous(range = c(1,5))
```


```{r}
col_brks = hist(alseis, plot=F)$breaks
col_indices = as.numeric(cut(alseis, col_brks))
cols = rev(terrain.colors(length(col_brks)))
plot(BCI_xy, cex=2, pch=19, col=cols[col_indices])
```
```{r}
# calculate Euclidean distance between abundance and spatial coordinates
ab_dist = dist(alseis)
xy_dist = dist(BCI_xy)
max_dist = max(xy_dist) / 2
```

```{r}
plot(xy_dist, ab_dist)
abline(lm(ab_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, ab_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)
```
```{r}
# compute correlation
obs_cor = cor(xy_dist, ab_dist)
obs_cor
```

```{r}
# carry out a permutation test for significance:
nperm = 1000
null_cor = obs_cor
for (i in 2:nperm) {
    # shuffle the rows of the spatial coordinates
    tmp_xy = BCI_xy[sample(nrow(BCI_xy)), ]
    # correlation between the shuffled spatial coordinates and ab_dist
    null_cor[i] = cor(dist(tmp_xy), ab_dist)
}
# compute the p-value
sum(null_cor >= obs_cor) / nperm 
```
```{r}
ab_mantel = mantel(xy_dist, ab_dist)
ab_mantel
```

```{r}
##examine community
## compute bray curtis distance for the community matrix
comm_dist = vegdist(BCI)
plot(xy_dist, comm_dist)
abline(lm(comm_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, comm_dist), lwd=3, col='pink')
lines(lowess(xy_dist, comm_dist, f=0.1), lwd=3, col='blue')

abline(v = max_dist, col='red', lwd=3, lty=2)
```
```{r}
comm_mantel = mantel(xy_dist, comm_dist)
comm_mantel
```
```{r}
ab_corlog = mantel.correlog(ab_dist, xy_dist)
comm_corlog = mantel.correlog(comm_dist, xy_dist)
ab_corlog
comm_corlog
##no significance in the abundance of Alseis but there is significant in the community structure (non linear)
```
```{r}
par(mfrow=c(1,2))
plot(ab_corlog)
mtext(side=3, 'Species Abundance')
abline(v = max_dist, col='red', lwd=3, lty=2)
plot(comm_corlog)
mtext(side=3, 'Community Composition')
abline(v = max_dist, col='red', lwd=3, lty=2)
```
Rare
```{r}
andira = BCI[,"Andira.inermis"]
andira
```
```{r}
plot(BCI_xy, cex = andira/max(andira))
```
```{r}
##ggplot
ggplot(BCI_xy, aes(x=x, y=y, size=andira)) + geom_point() + scale_size_continuous(range = c(1,5))
```

```{r}
col_brks = hist(andira, plot=F)$breaks
col_indices = as.numeric(cut(andira, col_brks))
cols = rev(terrain.colors(length(col_brks)))
plot(BCI_xy, cex=2, pch=19, col=cols[col_indices])
```
```{r}
andira_dist = dist(andira)
xy_dist = dist(BCI_xy)
max_dist = max(xy_dist) / 2
```
```{r}
plot(xy_dist, andira_dist)
abline(lm(andira_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, andira_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)
```
```{r}
andira_mantel = mantel(xy_dist, andira_dist)
andira_mantel
```
```{r}
andira_corlog = mantel.correlog(andira_dist, xy_dist)
andira_corlog
```
```{r}
par(mfrow=c(1,2))
plot(andira_corlog)
mtext(side=3, 'Species Abundance')
abline(v = max_dist, col='red', lwd=3, lty=2)
```
There is no evidence of spatial dependance in both the common (Alseis) and rare (Andira) species. 

2.one species
```{r}
sp_ids = c("Cordia.lasiocalyx", "Hirtella.triandra",
           "Picramnia.latifolia", "Quassia.amara",
           "Tabernaemontana.arborea", "Trattinnickia.aspera", 
           "Xylopia.macrantha")
BCI_ids = cbind("Drypetes.standleyi", "Cordia.lasiocalyx", "Hirtella.triandra",
           "Picramnia.latifolia", "Quassia.amara",
           "Tabernaemontana.arborea", "Trattinnickia.aspera", 
           "Xylopia.macrantha")
```

```{r}
sp_a <- apply(BCI_ids, 1, function(x) sum(x > 0))
sp_data = data.frame(sp_a, BCI, BCI_xy)
BCI_sub = subset(sp_data, select = BCI_ids)
```
```{r}
library(nlme)
sp_lm = gls(Drypetes.standleyi ~ Hirtella.triandra, data=BCI_sub)
plot(Variogram(sp_lm, form= ~ x + y))
```
```{r}
res = residuals(sp_lm)
plot(dist(sp_data[, c('x', 'y')]), dist(res))
lines(lowess(dist(sp_data[, c('x', 'y')]), dist(res)), col='red', lwd=2)
abline(v = max_dist, col='red', lwd=3, lty=2)
```
```{r}
sp_exp = update(sp_lm, corr=corExp(form=~x + y))
plot(Variogram(sp_exp, maxDist = max_dist))
```
```{r}
plot(Variogram(sp_exp, resType='normalized', maxDist = max_dist))
```
```{r}
sp_exp_nug = update(sp_exp, corr=corExp(form=~x + y, nugget=T))
plot(Variogram(sp_exp_nug, maxDist = max_dist))
```


```{r}
sp_rat_nug = update(sp_lm, corr=corRatio(form=~x + y, nugget=T))
plot(Variogram(sp_rat_nug, maxDist = max_dist))
plot(Variogram(sp_rat_nug, resType='n', maxDist = max_dist))
```
```{r}
col_brks = hist(residuals(sp_rat_nug), plot=F)$breaks
col_indices = as.numeric(cut(residuals(sp_rat_nug), col_brks))
cols = rev(terrain.colors(length(col_brks)))
plot(BCI_xy, cex=2, pch=19, col=cols[col_indices])
```


```{r}
anova(sp_lm, sp_exp, sp_exp_nug, sp_rat_nug, test=F)
summary(sp_lm)
summary(sp_exp)
summary(sp_exp_nug)
summary(sp_rat_nug)
```
Rational model with the nugget fit better then without spatial error terms

all species 
```{r}
sp_all_lm = gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra + Picramnia.latifolia + Quassia.amara + Tabernaemontana.arborea + Trattinnickia.aspera + Xylopia.macrantha , data=sp_data)
plot(Variogram(sp_all_lm, form= ~ x + y))
```
```{r}
res = residuals(sp_all_lm)
plot(dist(sp_data[, c('x', 'y')]), dist(res))
lines(lowess(dist(sp_data[, c('x', 'y')]), dist(res)), col='red', lwd=2)
abline(v = max_dist, col='red', lwd=3, lty=2)
```
```{r}
sp_all_exp = update(sp_all_lm, corr=corExp(form=~x + y))
plot(Variogram(sp_all_exp, maxDist = max_dist))
```
```{r}
plot(Variogram(sp_all_exp, resType='normalized', maxDist = max_dist))
```
```{r}
sp_all_exp_nug = update(sp_all_exp, corr=corExp(form=~x + y, nugget=T))
plot(Variogram(sp_all_exp_nug, maxDist = max_dist))
```


```{r}
sp_all_rat_nug = update(sp_all_lm, corr=corRatio(form=~x + y, nugget=T))
plot(Variogram(sp_all_rat_nug, maxDist = max_dist))
plot(Variogram(sp_all_rat_nug, resType='n', maxDist = max_dist))
```
```{r}
col_brks = hist(residuals(sp_all_rat_nug), plot=F)$breaks
col_indices = as.numeric(cut(residuals(sp_all_rat_nug), col_brks))
cols = rev(terrain.colors(length(col_brks)))
plot(BCI_xy, cex=2, pch=19, col=cols[col_indices])
```
```{r}
col_brks = hist(residuals(sp_all_exp_nug), plot=F)$breaks
col_indices = as.numeric(cut(residuals(sp_all_exp_nug), col_brks))
cols = rev(terrain.colors(length(col_brks)))
plot(BCI_xy, cex=2, pch=19, col=cols[col_indices])
```

```{r}
anova(sp_all_lm, sp_all_exp, sp_all_exp_nug, sp_all_rat_nug, test=F)
summary(sp_all_lm)
summary(sp_all_exp)
summary(sp_all_exp_nug)
summary(sp_all_rat_nug)
```
Spatial error terms did not effect the model


Adding spatial error terms (rational nugget) for the model with one species improved the model.  Fitting error to the model that included all species improved the model, but adding nuggets had no effect.  Overall, coefficients for both models had an impact, but this was larger for the model with only one species.  Spatial error terms can control for spatial dependence between neighbors and therefore improve our model. 


